---
title: "Extract features for CART4 + PT1Y9 datasets"
output:
  pdf_document: default
  html_notebook: default
---

```{r, warning=F, message=F}
library(Seurat)
library(tidyverse)
library(matrixStats)
theme_set(theme_bw())
```

# Import CART4 dataset

```{r}
data.dir <- "CART4/filtered_feature_bc_matrix/"
list.files(data.dir) # Should show barcodes.tsv.gz, features.tsv.gz, and matrix.mtx.gz
data <- Read10X(data.dir)
sobj <-  CreateSeuratObject(counts = data$`Gene Expression`)
sobj[['Protein']] <- CreateAssayObject(counts = data$`Antibody Capture`)
```

```{r}
sobj@assays
```

For CART4 dataset:

* 4807 cells, 33.5k genes, 7 proteins
* Output should have 4807 cells

# Extract features

Define `nProtein` as the number of protein features (e.g. 7). Define `nRNA` as the number of gene features (e.g. 33.5k).

* `nCount_Protein`: sum of protein counts
* `nFeature_Protein`: number of non-zero gene counts
* `percFeature_Protein`: `nFeature_Protein` / `nProtein` 
* `average_Protein`: mean of protein counts 
* `median_Protein`: median of protein counts 
* `variance_Protein`: variance of protein counts 
* `nCount_RNA`: sum of gene counts
* `nFeature_RNA`: number of non-zero gene counts
* `percFeature_RNA`: `nFeature_RNA` / `nRNA` 
* `average_RNA`: mean of gene counts 

```{r}
out.df <- sobj@meta.data %>% select(-orig.ident)

# remove sparsity to do calculations
protein_matrix <- as.matrix(sobj@assays$Protein@counts)
RNA_matrix <- as.matrix(sobj@assays$RNA@counts)

# store number of protein and gene features
nProtein <- dim(protein_matrix)[1]
nRNA <- dim(RNA_matrix)[1]

out.df <- out.df %>% mutate(
  percFeature_Protein = nFeature_Protein / nProtein,
  average_Protein = colMeans(protein_matrix),
  median_Protein = colMedians(protein_matrix),
  variance_Protein= colVars(protein_matrix),
  percFeature_RNA= nFeature_Protein / nRNA,
  average_RNA= colMeans(RNA_matrix)
)
```

# Write ouput

```{r}
# out.df %>% rownames_to_column("cell_ID") %>%
#   write.table("2021-04-13_CART4_features_4807x10.txt", sep="\t", quote=F, row.names=F)
```

# Examine distributions of extracted features

```{r}
# pdf("2021-04-13_CART4_extracted_features_histogram.pdf", width=10, height=5)
pivot_longer(out.df, everything(), names_to="feature") %>%
  ggplot(aes(value)) + geom_histogram(bins=50) +
  theme(axis.text.x = element_text(angle=45)) +
  facet_wrap(~feature, ncol=5, scales="free")
# dev.off()
```


Other potential features that we could implement:

* Avg distance to nearest 1, 5, 10, 15… RNA neighbors 
* Avg distance to nearest 1, 5, 10, 15… protein neighbors 
* Cluster identity

# Repeat for PT1Y9 dataset

```{r}
sobj2 <- readRDS("PT1Y9_import.RDS")
sobj2@assays
```

* 3072 cells x 16k genes and 21 proteins

```{r}
out.df2 <- sobj2@meta.data %>% dplyr::select(-orig.ident)

# remove sparsity to do calculations
protein_matrix <- as.matrix(sobj2@assays$ADT@counts)
RNA_matrix <- as.matrix(sobj2@assays$RNA@counts)

# store number of protein and gene features
nProtein <- dim(protein_matrix)[1]
nRNA <- dim(RNA_matrix)[1]

out.df2 <- out.df2 %>% mutate(
  nCount_Protein = nCount_ADT,
  nFeature_Protein = nFeature_ADT,
  percFeature_Protein = nFeature_ADT / nProtein,
  average_Protein = colMeans(protein_matrix),
  median_Protein = colMedians(protein_matrix),
  variance_Protein= colVars(protein_matrix),
  percFeature_RNA= nFeature_ADT / nRNA,
  average_RNA= colMeans(RNA_matrix)
)

out.df2 <- out.df2 %>% dplyr::select(-contains("ADT")) %>%
  dplyr::select(-percent.mt)

# out.df2 %>% rownames_to_column("cell_ID") %>%
#   write.table("2021-04-14_PT1Y9_features_3072x10.txt", sep="\t", quote=F, row.names=F)
```

```{r}
# pdf("2021-04-14_PT1Y9_extracted_features_histogram.pdf", width=10, height=5)
pivot_longer(out.df2, everything(), names_to="feature") %>%
  ggplot(aes(value)) + geom_histogram(bins=50) +
  theme(axis.text.x = element_text(angle=45)) +
  facet_wrap(~feature, ncol=5, scales="free")
# dev.off()
```

